---
title: "Returns of financial stocks"
author: "Akos"
date: "9/17/2021"
output: html_document
slug: blog2
categories:
- ''
- ''
---

```{r load-libraries, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(fivethirtyeight)
library(here)
library(skimr)
library(janitor)
library(vroom)
library(tidyquant)
library(rvest) # to scrape wikipedia page
library(scales) # to make chart scales "dynamic"
library(ggrepel) # necessary for the inflation chart
```

```{r load_nyse_data, message=FALSE, warning=FALSE}
nyse <- read_csv(here::here("data","nyse.csv"))
```

Based on this dataset, create a table and a bar plot that shows the number of companies per sector, in descending order

```{r companies_per_sector}

nyse_sector<-nyse%>%
  group_by(sector) %>%
  summarise(number = n())%>%
  arrange(desc(number))
nyse_sector

nyse_sector<-data.frame(nyse_sector)

nyse_sector %>%
    ggplot(aes(x = number,y= reorder(sector, number)))+ geom_col() +
    labs(title = "Number of companies per sector (Desc)", x = "Number of companies", y = "Sectors")

```


Next, let's choose the [Dow Jones Industrial Aveareg (DJIA)](https://en.wikipedia.org/wiki/Dow_Jones_Industrial_Average) stocks and their ticker symbols and download some data. Besides the thirty stocks that make up the DJIA, we will also add `SPY` which is an SP500 ETF (Exchange Traded Fund). The following code 


```{r, tickers_from_wikipedia}

djia_url <- "https://en.wikipedia.org/wiki/Dow_Jones_Industrial_Average"


#get tables that exist on URL
tables <- djia_url %>% 
  read_html() %>% 
  html_nodes(css="table")


# parse HTML tables into a dataframe called djia. 
# Use purr::map() to create a list of all tables in URL
djia <- map(tables, . %>% 
               html_table(fill=TRUE)%>% 
               clean_names())


# constituents
table1 <- djia[[2]] %>% # the second table on the page contains the ticker symbols
  mutate(date_added = ymd(date_added),
         
         # if a stock is listed on NYSE, its symbol is, e.g., NYSE: MMM
         # We will get prices from yahoo finance which requires just the ticker
         
         # if symbol contains "NYSE*", the * being a wildcard
         # then we jsut drop the first 6 characters in that string
         ticker = ifelse(str_detect(symbol, "NYSE*"),
                          str_sub(symbol,7,11),
                          symbol)
         )

# we need a vector of strings with just the 30 tickers + SPY
tickers <- table1 %>% 
  select(ticker) %>% 
  pull() %>% # pull() gets them as a sting of characters
  c("SPY") # and lets us add SPY, the SP500 ETF

```

Now let us downlaod prices for all 30 DJIA consituents and the SPY ETF that tracks SP500 since January 1, 2020


```{r get_price_data, message=FALSE, warning=FALSE, cache=TRUE}
# Notice the cache=TRUE argument in the chunk options. Because getting data is time consuming, # cache=TRUE means that once it downloads data, the chunk will not run again next time you knit your Rmd

myStocks <- tickers %>% 
  tq_get(get  = "stock.prices",
         from = "2000-01-01",
         to   = Sys.Date()) %>% # Sys.Date() returns today's price
  group_by(symbol) 

glimpse(myStocks) # examine the structure of the resulting data frame
```


```{r calculate_returns, message=FALSE, warning=FALSE, cache=TRUE}
#calculate daily returns
myStocks_returns_daily <- myStocks %>%
  tq_transmute(select     = adjusted, 
               mutate_fun = periodReturn, 
               period     = "daily", 
               type       = "log",
               col_rename = "daily_returns",
               cols = c(nested.col))  

#calculate monthly  returns
myStocks_returns_monthly <- myStocks %>%
  tq_transmute(select     = adjusted, 
               mutate_fun = periodReturn, 
               period     = "monthly", 
               type       = "arithmetic",
               col_rename = "monthly_returns",
               cols = c(nested.col)) 

#calculate yearly returns
myStocks_returns_annual <- myStocks %>%
  group_by(symbol) %>%
  tq_transmute(select     = adjusted, 
               mutate_fun = periodReturn, 
               period     = "yearly", 
               type       = "arithmetic",
               col_rename = "yearly_returns",
               cols = c(nested.col))
```

Create a table where you summarise monthly returns for each of the stocks and `SPY`; min, max, median, mean, SD.

```{r summarise_monthly_returns}

summary_monthly_returns <- myStocks_returns_monthly %>% 
  summarise(min = min(monthly_returns), max = max(monthly_returns), median = median(monthly_returns), mean = mean(monthly_returns), SD = STDEV(monthly_returns)) %>% 
  arrange(desc(SD))
summary_monthly_returns
```


Plot a density plot, using `geom_density()`, for each of the stocks

```{r density_monthly_returns}

myStocks_returns_monthly %>% 
  ggplot(aes(x = monthly_returns, fill = symbol)) + geom_density() + facet_grid(rows = vars(symbol)) + labs(title = "Density Plot of Stock's Monthly Return", x = "Monthly Returns", y ="Density")

myStocks_returns_monthly %>% 
  ggplot(aes(x = monthly_returns, fill = symbol)) + 
  geom_density() + 
  facet_grid(rows = vars(symbol)) +  # Grid to better compare each stock's riskiness
  labs(title = "Density Plot of Stock's Monthly Return", x = "Monthly Returns", y ="Density") + 
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) + 
  coord_cartesian(xlim=c(-0.2,0.2)) +
  NULL


```

What can you infer from this plot? Which stock is the riskiest? The least risky? 

> 

From this plot we can see that the stocks present different riskiness. For riskier stocks the monthly returns are more spread out, while for less risky stocks the monthly returns converge more towards one specific return, which results in a higher curve. According to this, we can see that the riskiest stock is APPL and the least risky is SPY (we can double check this by looking at the previous table, which we arranged by Standard deviation since this measures how spread the data is). Moreover, SPY is an index for S&P500, so this is coherent with it being the least risky of all stocks.


Finally, make a plot that shows the expected monthly return (mean) of a stock on the Y axis and the risk (standard deviation) in the X-axis. Please use `ggrepel::geom_text_repel()` to label each stock

```{r risk_return_plot}

summary_monthly_returns %>% 
  ggplot(aes(x = SD, y= mean, label=symbol, color=symbol)) + 
  geom_point() + 
  geom_text_repel() + 
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) + 
  labs(title="Expected monthly return according to risk", x= "Risk", y="Expected monthly return") + 
  NULL
```

What can you infer from this plot? Are there any stocks which, while being riskier, do not have a higher expected return?

> 

We can see that some stocks are among the riskiest stocks, however their expected monthly return is the same as stocks which are more than twice less risky. This is the case especially for INTC and CSCO.

